Convergence rate of MAP estimates \\ for the exponential family

### 2021-01-26 main.tex
Try fixing UTF8 error
Fix UTF8 error
Use semi colon instead of || in all divergences

\paragraph{Conjugate Prior}
The conjugate prior for $p(X|\natp)$ is
\begin{align}
    \pi(\natp) &\propto \exp( - n_0 \bregman(\natp || \natp_0) )
\end{align}
where $n_0$ and $\natp_0$ are (hyper)parameters of the prior.
Intuitively, $n_0$ is a number of fictive points observed from a distribution with parameter $\natp_0$.
Finally, $\bregman(\natp || \natp_0)$ is the Bregman divergence induced by $\logpart$ between $\natp$ and $\natp_0$
\begin{align}
    \bregman (\natp || \natp_0)
    & = \logpart(\natp) - \logpart(\natp_0) 
    - \langle \nabla \logpart(\natp_0)  , \natp - \natp_0 \rangle
\end{align}
with $\nabla \logpart(\natp_0) = \expect[\natp_0]{T(X)} =: \meanp_0$ the mean parameter associated to $\natp_0$.

\paragraph{MAP.}
The negative log-likelihood of the prior is then
\begin{align*}
    -\log \pi(\natp) = n_0 (\logpart(\natp)  - \natp^\top \meanp_0 ) + \cst
\end{align*}
Thus the joint NLL of $(x_1,\dots,x_n,\natp)$ is
\begin{align}
    -\log p(X|\natp)\pi(\natp) 
    = (n_0+n) \logpart (\natp) 
    - \theta^\top \left(n_0 \meanp_0 + \sum_{i=1}^n T(x_i) \right) + \cst \; .
\end{align}
Minimizing this expression over $\natp$ yields the Maximum A Posteriori estimate
\begin{align}
    \hat \natp = \argmin_\natp -\log p(X|\natp) + n_0 \bregman(\natp || \natp_0)
\end{align}
such that the MAP is
\begin{align}
    \nabla \logpart(\hat \natp) = \hat \meanp
    = \frac{n_0 \meanp_0 + \sum_{i=1}^n T(X_i) }{n_0+n} \; .
\end{align}
The MAP estimate is a random quantity. We wish to bound its deviation the optimum $\natp^*$ or $\meanp^*$.

\section{Straightforward Convergence Rate}
In the realizable case, the suboptimality on the population log-likelihood is exactly the KL between our current model and the true distribution
\begin{align}
    &\expect[X\sim p(.|\natp^*)]{-\log p(X|\natp) + \log p(X|\natp^*)}\\
    =& \KL( p(.|\natp^*) || p(.|\natp)) \\
    =& \bregman (\natp || \natp^*) \\
    =& \bregmanconj ( \meanp^* || \meanp) \\
    =& \conj(\mu^*) + \logpart(\natp) - \langle \natp , \mu^* \rangle
    %& = \logpart(\natp) - \logpart(\natp^*) - \meanp^{*\top}(\natp - \natp^*) \; .
\end{align}
where $\conj$ is the entropy, the convex conjugate of the log-partition.
The relationship between Bregman divergences and Fenchel conjugacy is well explained in \citet{wainwright2008graphical}, and \citet{agarwal2010geometric}. 
The question is: how does this quantity behave when $\natp$ is the maximum-likelihood or the MAP estimate ? Can we get bounds -- in expectation or high-probability -- on the following quantities ?
\begin{align}
	\label{eq:bregmanMLE}
	\bregmanconj \left (\E [T(X)] ;  \inv{n}  \smallsum_i T(X_i) \right ) \; , \\
	\label{eq:bregmanMAP}
	\bregmanconj \left (\E [T(X)] ; \frac{n_0 \mu_0 + \smallsum_i T(X_i)}{n_0+n} \right ) \; .
\end{align}

\paragraph{If $\conj$ is $L$-Lipschitz} (e.g. $\logpart$ is defined within the $\ell^2$-ball of radius $L$), then
\begin{align}
    \bregmanconj(\mu^* || \mu) 
    \leq L \norm{\mu^* - \mu} + \norm{\natp} \norm{\mu^* - \mu}
    \leq 2L \norm{\mu^* - \mu}
\end{align}
so $\bregmanconj$ is $2L$-Lipschitz.
Since the empirical average converges in expectation to the population mean at a rate of $1/\sqrt{n}$ in $\ell^2$ norm, we know that this bound applies to the log-likelihood.

\paragraph{If $\conj$ is $L$-smooth} (e.g. $\logpart$ is $L^{-1}$-strongly convex), then
\begin{align}
    \bregmanconj(\mu^* || \mu) 
    \leq \frac{L}{2} \norm{\mu^* - \mu}^2
\end{align}
so $\bregmanconj$ is upper bounded by a quadratic. In expectation, it should converge at a rate $1/n$.

@@

In the limit, the MAP estimates reach the Cramer-Rao lower bound. This can be seen by approximating the Bregman divergence using a second order Taylor expansion
\begin{align}
    \bregmanconj(\mu^* || \mu) 
    &= \half (\mu - \mu^*)^\top \nabla^2\conj(\mu^*)(\mu - \mu^*)  
    + O(\norm{\mu - \mu^*}^3)\\
    &=\half  \norm{\mu^* - \mu}^2_{\nabla^2\conj(\mu^*)}
    + O(\norm{\mu - \mu^*}^3)
\end{align}
where we introduced the Mahalanobis distance induced by the matrix 
\begin{align}
    \nabla^2\conj(\mu^*) 
    = \nabla^2\logpart(\natp^*)^{-1} 
    = \Cov_{\natp^*}[T(X)]^{-1} 
    =: \mSigma^{-1}  \; .
\end{align}
To exploit this approximation, let us extend the $\ell^2$ norm results to a Mahalanobis distance induced by matrix $\mM$
\begin{align}
	\E \norm{\mu^* -  \inv{n}  \smallsum_i T(x_i)}_\mM^2 = \inv{n} \Tr(\mM \mSigma)
\end{align}
where $\mSigma$ is the covariance of $T(X)$. We retrieve $\nicefrac{\sigma^2}{n}$, the variance divided by the number of samples, when the metric is the identity $\mM=\mI$.  For the MLE we get
\begin{align}
	\E \bregmanconj \left (\E [T(X)] ;  \inv{n}  \smallsum_i T(X_i) \right ) = \frac{d}{2n} + O(n^{- \frac{3}{2}})
\end{align}
\RLP{Maybe that is Cramer-Rao, but check again.} And for the MAP we  get 
\begin{align}
	\E \norm{\mu^* -  \frac{n_0 \mu_0 + \smallsum_i T(x_i)}{n_0+n} }^2_{\mSigma^{-1}}
	&= \frac{n d}{(n+n_0)^2}  +  \frac{n_0^2}{(n+n_0)^2} \norm{\mu^* -  \mu_0}^2_{\mSigma^{-1}} \\
	&= O\left(\frac{d}{n} \right) + O\left(\frac{\norm{\mu^* -  \mu_0}^2_{\mSigma^{-1}} }{n^2} \right) \; .
\end{align}
Remark how the variance does not even appear, only the dimension divided by $n$ really matters. I find this result quite spectacular : the convergence speed of MLE is not affected by the covariance of sufficient statistics, and for MAP it matters only in the $O(n^{-2})$ term.
Anyway that's all good for asymptotic results, but we are interested in a finite sample analysis. How do we get this ?

\section{Stochastic Bregman Proximal Point}
We see that the MAP estimate minimizes the sum of a stochastic loss $ -\log p(X|\natp)$ and a deterministic divergence to an initial point $n_0 \bregman(\natp || \natp_0)$. This is a stochastic Bregman proximal step with step-size $\inv{n_0}$. This can also be seen at each step since
\begin{align}
    \hat \natp_{n+1} 
    &= \argmin_\natp -\log p(x_{n}|\natp) + (n_0 + n) \bregman(\natp || \hat \natp_n) \\
    &= \argmin_\natp f(\natp; x_{n}) +  \inv{\gamma_{n}}\bregman(\natp || \hat \natp_n) \; .
\end{align}
Hence the MAP estimate can also be seen as the result of a stochastic proximal Bregman point algorithm with step-size $\gamma_n = \inv{n_0 + n}$ at step $n$.

@@

\section{Stochastic Mirror Descent (SMD)}
\subsection{MAP as SMD}
Let $f(\natp) := \expect[x]{-\log p(x|\natp)} = - \langle \mu_*, \natp\rangle + \logpart(\natp)$. In words, $f$ is linear modification of a convex function $\logpart$, which we can access only through noisy estimates of $\mu_*$. It turns out that the MAP estimate can also be seen as the iterates of stochastic mirror descent with mirror map $\logpart$. First let's recall mirror descent iteration
\begin{align}
	\natp_{t+1} 
	&:= \argmin_\natp \lr \linear_f(\natp; \natp_t) + \bregman(\natp || \natp_t)  \\
	& = \nabla \conj (\nabla \logpart(\natp_t) - \lr \nabla f(\natp_t))
\end{align}
where $\linear_f(\natp; \natp_t) = f(\natp_t) + \langle \nabla f(\natp_t), \natp - \natp_t \rangle$ is the linear approximation of $f$ in $\natp_t$ evaluated at $\natp$. Solving this problem require solving problems of the form $\argmin_\natp - \langle c, \natp \rangle + \logpart(\natp)$, eg computing the convex conjugate of $\logpart$. 
Note that finding $\natp_*$ is done with 1 step of mirror descent. Indeed plugging in definitions of $f$ and $\mu$ yields
\begin{align}
	&\mu_{t+1}  = \mu_t - \lr (\mu_t - \mu_*) \\
	\implies &\mu_t  = \mu_* + (1- \lr)^t (\mu_0 - \mu_*)
\end{align}
which shows exponential convergence and 1-step convergence when $\lr =1$. Back to our sheep, the MAP iteration can be cast as stochastic mirror descent (SMD),  with $g_t = \nabla f(\natp_t, x_{t+1})$ a stochastic estimate  of $\nabla f(\natp_t)$
\begin{align}
    \hat \natp_{n+1} 
    &= \argmin_\natp 
    - \langle T(x_{t+1}), \natp \rangle  + \logpart(\natp) + (n_0 + n) \bregman(\natp || \natp_n) \\
    %&(n_0+n)\left(\logpart(\natp) - \logpart(\natp_n) - \langle \nabla \logpart(\natp_n)  , \natp - \natp_n \rangle \right)  \\
    &= \argmin_\natp - \langle T(x_{t+1}), \natp \rangle   + \logpart(\natp_n) + \langle \nabla \logpart(\natp_n)  , \natp - \natp_n \rangle + (n_0+n + 1) \bregman(\natp || \natp_n) \\
    &= \argmin_\natp \linear_f(\natp;\natp_n, x_{t+1}) + (n_0 + n + 1) \bregman(\natp || \natp_n)
\end{align}
where $\linear_f(\natp;\natp_n, x_{t+1})$ is the stochastic linearization of $f$ at $\natp_n$ evaluated at $\natp$ with randomness coming from $x_{t+1}$. This is the formula for stochastic mirror descent (SMD) applied to $f$ with mirror map $\logpart$ and step-size $\gamma_n = \inv{n_0 + n + 1}$.

\subsection{Relative Smoothness for Mirror Descent}
In the classic setting, SMD is studied under strong-convexity assumption on the mirror map $\logpart$ \citep{bubeck2015convex}.
In our setting this is not always true -- eg gaussians.
However a recent and fast-expanding body of work is concerned with a new assumption: relative smoothness and relative strong-convexity.
\begin{align}
	& \text{$f$ is $L$-smooth relative to $h$} \\
	\iff & Lh - f \ \text{convex} \\
	\iff & f(y) \leq 	f(x) + \langle \nabla f(x) , y-x \rangle + L\cB_h(y||x), \forall x,y \\
	\iff & \cB_f(y||x) \leq L \cB_h(y||x), \forall x,y \\
	\iff & \nabla^2 f(x) \leq L \nabla^2 h(x),  \forall x
\end{align} 
where the last equivalence holds only when $f$ and $h$ are twice differentiable. In words, $f$ is upper bounded by it linear approximation plus the $h$-Bregman divergence, which can also be seen as a bound between divergences, or more locally as a bound between Hessians. Similarly, $f$ is $\mu$ strongly-convex relative to $h$ if
\begin{align}
	& f-\mu h \ \text{ convex} \\
	\iff & f(x) + \langle \nabla f(x) , y-x \rangle + \mu \cB_h(y||x) \leq f(y)\forall x,y \\
	\iff & \mu\cB_h(y||x) \leq \cB_f(y||x) \forall x,y \\
	\iff & \mu \nabla^2 h(x) \leq \nabla^2 f(x)  \forall x \; .
\end{align}

Another way to view this elegant generalization of smoothness and strong-convexity is as a transfer of the Loewner partial order on symmetric matrices to functions, via the Hessian. 
As such it can be applied to many functions that were out of reach for $\ell^2$ norm, by taking the appropriate reference function.  For instance $h(x) = -\log(x)$ or $h(x) = x^4$. 
As early as 2011, \citet{birnbaum2011distributed} showed $O(\inv{t})$ convergence rate for mirror descent under smoothness assumption relative to the mirror map. 
More precisely, he proved that when $f$ is $L$-smooth relative to $h$, then the suboptimality of the sequence 
\begin{align}
	x_{t+1}  = \argmin_x \langle \nabla f(x_t), x \rangle + \cB_h(x || x_t)
\end{align}
is upper bounded by the simple formula
\begin{align}
	\implies f(x_t) - f(x_*) \leq \frac{L \cB_h(x_* || x_0) }{t} \; .
\end{align}
These notions were rediscovered and expanded by \citet{bauschke2017descent} and \citet{lu2018relatively}. If you need to read one, pick \citet{lu2018relatively} -- I found it much much easier and more enjoyable to read.  This latter paper also derived a linear convergence rate for mirror descent under relative smoothness and strong-convexity, with the relative condition number $\frac{L}{\mu}$ appearing.

@@

	\expect[X_{1\dots n}]{\bregmanconj ( \mu_* ||\hat \mu) }

@@

Our functions of interest is $f(\natp)= \bregman (\natp || \natp^*)= \bregmanconj ( \meanp^* || \meanp) = g(\meanp)$, with minimum $f^*=0$. 
If $\logpart$ is self-concordant, then so is $f$, but not necessarily $g$.
The gradient and Hessian of $f$ are
\begin{align}
    f(\natp) 
    &= \logpart(\natp) - \logpart(\natp^*) - \langle \meanp^*, \natp - \natp^* \rangle \\
    \nabla f(\natp) 
    &= \meanp - \meanp^* = \expect[p(X|\natp)]{T(X)} - \expect[p(X|\natp^*)]{T(X)} \\
    \nabla^2 f(\natp) 
    &= \mSigma(\natp) = \Cov_{p(X|\natp)}[T(X)]
\end{align}
so that we get the bound.
\begin{align}
    \bregmanconj(\mu^* || \mu) 
    \leq \decrement(\natp)^2 
    = \norm{\mu^* - \mu(\theta)}^2_{\mSigma(\theta)^{-1}} 
    \leq 0.46
\end{align}
% If we were to use $g(\meanp)$ instead, we would get dirty third order derivatives
% \begin{align}
%     g(\meanp) 
%     &= \conj(\meanp^*) - \conj(\meanp) - \langle \natp, \meanp^* - \meanp \rangle \\
%     \nabla g(\meanp) 
%     &= -\natp - \nabla^2 \conj(\meanp) (\meanp^* - \meanp) + \natp \\
%     &= \nabla^2 \conj(\meanp) (\meanp - \meanp^*) \\
%     \nabla^2 g(\meanp) 
%     &= \nabla^2 \conj(\meanp) + \nabla^3 \conj(\meanp ) (\meanp - \meanp^*)
% \end{align}
% where the last expression involves a third order tensor.

Finally, if instead we were looking at a different function switching the role of $\meanp$ and $\meanp^*$,  $h(\meanp) = \bregmanconj ( \meanp || \meanp^*)$, then we would get
\begin{align}
    \nabla h(\meanp ) 
    &= \natp - \natp^* \\
    \nabla^2 h(\meanp ) 
    &= \nabla^2 \conj(\meanp) 
    = \nabla^2 \logpart(\natp) ^{-1}
    = \Cov_{p(X|\natp)}[T(X)]^{-1}\\
    \implies 
    \decrement(\meanp)
    &= \Var_{p(X|\natp)}[(\natp - \natp^*)^T T(X)] \; .
\end{align}
This is just a remark. I don't think it can help us to get anywhere.

@@

Proofs of convergence for mirror descent with relative smoothness rely on a specific property of Bregman divergences, familiar to information geometry folks. If $x_+$ is solution to 
\begin{align}
	\min	_{x\in C} \overbrace{f(x) + \cB_h(x || y)}^{\phi(x)}
\end{align}
where $C$ is a closed convex set, $f$ is a convex function, $\cB_h$ is the Bregman divergence induce by some convex function $h$ (we will drop the index, and use a semicolon instead of double bars to lighten the notation), and $y$ is some reference vector. Then 
\begin{align}
	\forall x, f(x) + \cB(x ;y) \geq f(x_+) + \cB(x ; x_+) +\cB(x_+ ; y) \; .
\end{align}
This property is an analog of the Pythagorean theorem for generalized projections (setting $f=0$ and $h=\|.\|^2$).

>>>>

\paragraph{Conjugate Prior}
The conjugate prior for $p(X|\natp)$ is
\begin{align}
    \pi(\natp) &\propto \exp( - n_0 \bregman(\natp ; \natp_0) )
\end{align}
where $n_0$ and $\natp_0$ are (hyper)parameters of the prior.
Intuitively, $n_0$ is a number of fictive points observed from a distribution with parameter $\natp_0$.
Finally, $\bregman(\natp ; \natp_0)$ is the Bregman divergence induced by $\logpart$ between $\natp$ and $\natp_0$
\begin{align}
    \bregman (\natp ; \natp_0)
    & = \logpart(\natp) - \logpart(\natp_0) 
    - \langle \nabla \logpart(\natp_0)  , \natp - \natp_0 \rangle
\end{align}
with $\nabla \logpart(\natp_0) = \expect[\natp_0]{T(X)} =: \meanp_0$ the mean parameter associated to $\natp_0$.

\paragraph{MAP.}
The negative log-likelihood of the prior is then
\begin{align*}
    -\log \pi(\natp) = n_0 (\logpart(\natp)  - \natp^\top \meanp_0 ) + \cst
\end{align*}
Thus the joint NLL of $(x_1,\dots,x_n,\natp)$ is
\begin{align}
    -\log p(X|\natp)\pi(\natp) 
    = (n_0+n) \logpart (\natp) 
    - \theta^\top \left(n_0 \meanp_0 + \sum_{i=1}^n T(x_i) \right) + \cst \; .
\end{align}
Minimizing this expression over $\natp$ yields the Maximum A Posteriori estimate
\begin{align}
    \hat \natp = \argmin_\natp -\log p(X|\natp) + n_0 \bregman(\natp ; \natp_0)
\end{align}
such that the MAP is
\begin{align}
    \nabla \logpart(\hat \natp) = \hat \meanp
    = \frac{n_0 \meanp_0 + \sum_{i=1}^n T(X_i) }{n_0+n} \; .
\end{align}
The MAP estimate is a random quantity. We wish to bound its deviation the optimum $\natp^*$ or $\meanp^*$.

\section{Straightforward Convergence Rate}
In the realizable case, the suboptimality on the population log-likelihood is exactly the KL between our current model and the true distribution
\begin{align}
    &\expect[X\sim p(.|\natp^*)]{-\log p(X|\natp) + \log p(X|\natp^*)}\\
    =& \KL( p(.|\natp^*) ; p(.|\natp)) \\
    =& \bregman (\natp ; \natp^*) \\
    =& \bregmanconj ( \meanp^* ; \meanp) \\
    =& \conj(\mu^*) + \logpart(\natp) - \langle \natp , \mu^* \rangle
    %& = \logpart(\natp) - \logpart(\natp^*) - \meanp^{*\top}(\natp - \natp^*) \; .
\end{align}
where $\conj$ is the entropy, the convex conjugate of the log-partition.
The relationship between Bregman divergences and Fenchel conjugacy is well explained in \citet{wainwright2008graphical}, and \citet{agarwal2010geometric}. 
The question is: how does this quantity behave when $\natp$ is the maximum-likelihood or the MAP estimate ? Can we get bounds -- in expectation or high-probability -- on the following quantities ?
\begin{align}
	\label{eq:bregmanMLE}
	\bregmanconj \left (\E [T(X)] ;  \inv{n}  \smallsum_i T(X_i) \right ) \; , \\
	\label{eq:bregmanMAP}
	\bregmanconj \left (\E [T(X)] ; \frac{n_0 \mu_0 + \smallsum_i T(X_i)}{n_0+n} \right ) \; .
\end{align}

\paragraph{If $\conj$ is $L$-Lipschitz} (e.g. $\logpart$ is defined within the $\ell^2$-ball of radius $L$), then
\begin{align}
    \bregmanconj(\mu^* ; \mu) 
    \leq L \norm{\mu^* - \mu} + \norm{\natp} \norm{\mu^* - \mu}
    \leq 2L \norm{\mu^* - \mu}
\end{align}
so $\bregmanconj$ is $2L$-Lipschitz.
Since the empirical average converges in expectation to the population mean at a rate of $1/\sqrt{n}$ in $\ell^2$ norm, we know that this bound applies to the log-likelihood.

\paragraph{If $\conj$ is $L$-smooth} (e.g. $\logpart$ is $L^{-1}$-strongly convex), then
\begin{align}
    \bregmanconj(\mu^* ; \mu) 
    \leq \frac{L}{2} \norm{\mu^* - \mu}^2
\end{align}
so $\bregmanconj$ is upper bounded by a quadratic. In expectation, it should converge at a rate $1/n$.

@@

In the limit, the MAP estimates reach the Cramer-Rao lower bound. This can be seen by approximating the Bregman divergence using a second order Taylor expansion
\begin{align}
    \bregmanconj(\mu^* ; \mu) 
    &= \half (\mu - \mu^*)^\top \nabla^2\conj(\mu^*)(\mu - \mu^*)  
    + O(\norm{\mu - \mu^*}^3)\\
    &=\half  \norm{\mu^* - \mu}^2_{\nabla^2\conj(\mu^*)}
    + O(\norm{\mu - \mu^*}^3)
\end{align}
where we introduced the Mahalanobis distance induced by the matrix 
\begin{align}
    \nabla^2\conj(\mu^*) 
    = \nabla^2\logpart(\natp^*)^{-1} 
    = \Cov_{\natp^*}[T(X)]^{-1} 
    =: \mSigma^{-1}  \; .
\end{align}
To exploit this approximation, let us extend the $\ell^2$ norm results to a Mahalanobis distance induced by matrix $\mM$
\begin{align}
	\E \norm{\mu^* -  \inv{n}  \smallsum_i T(x_i)}_\mM^2 = \inv{n} \Tr(\mM \mSigma)
\end{align}
where $\mSigma$ is the covariance of $T(X)$. We retrieve $\nicefrac{\sigma^2}{n}$, the variance divided by the number of samples, when the metric is the identity $\mM=\mI$.  For the MLE we get
\begin{align}
	\E \bregmanconj \left (\E [T(X)] ;  \inv{n}  \smallsum_i T(X_i) \right ) = \frac{d}{2n} + O(n^{- \frac{3}{2}})
\end{align}
\RLP{Maybe that is Cramer-Rao, but check again.} And for the MAP we  get 
\begin{align}
	\E \norm{\mu^* -  \frac{n_0 \mu_0 + \smallsum_i T(x_i)}{n_0+n} }^2_{\mSigma^{-1}}
	&= \frac{n d}{(n+n_0)^2}  +  \frac{n_0^2}{(n+n_0)^2} \norm{\mu^* -  \mu_0}^2_{\mSigma^{-1}} \\
	&= O\left(\frac{d}{n} \right) + O\left(\frac{\norm{\mu^* -  \mu_0}^2_{\mSigma^{-1}} }{n^2} \right) \; .
\end{align}
Remark how the variance does not even appear, only the dimension divided by $n$ really matters. I find this result quite spectacular : the convergence speed of MLE is not affected by the covariance of sufficient statistics, and for MAP it matters only in the $O(n^{-2})$ term.
Anyway that's all good for asymptotic results, but we are interested in a finite sample analysis. How do we get this ?

\subsection{Posterior Expected Loss}

A conjugate prior for the exponential family with sufficient statistic $T(X)$ and natural parameter $\natp$ is 
\begin{align}
    \pi(\natp) 
    \propto \exp(-n_0 \bregman(\natp ; \natp_0)) 
    \propto \exp(-n_0 \logpart(\natp) + \langle n_0 \mu_0 ; \natp \rangle)
\end{align}
so we see that $\pi$ is part of the exponential family, with sufficient statistic $(\natp, \logpart(\natp))$ and natural parameters $(n_0 \mu_0 , -n_0)$. When we update this prior with Bayes formula on dataset $\cD=(x_1,\dots,x_n)$, we get a posterior $\pi(\natp | \cD)$ with natural parameters $(n_0 \mu_0 + \sum_i T(x_i) , -(n_0 + n))$. This is well detailed in \citet{agarwal2010geometric}.

Given this posterior, a natural quantity to consider is the expected loss. Thus we might ask "what is the expectation of the posterior expected suboptimality?"
\begin{align}
    \expect[\cD\sim\natp_*]{\expect[\pi(\natp|\cD)]{\bregman(\natp ; \natp_*)}}
\end{align}

\section{Stochastic Bregman Proximal Point}
We see that the MAP estimate minimizes the sum of a stochastic loss $ -\log p(X|\natp)$ and a deterministic divergence to an initial point $n_0 \bregman(\natp ; \natp_0)$. This is a stochastic Bregman proximal step with step-size $\inv{n_0}$. This can also be seen at each step since
\begin{align}
    \hat \natp_{n+1} 
    &= \argmin_\natp -\log p(x_{n}|\natp) + (n_0 + n) \bregman(\natp ; \hat \natp_n) \\
    &= \argmin_\natp f(\natp; x_{n}) +  \inv{\gamma_{n}}\bregman(\natp ; \hat \natp_n) \; .
\end{align}
Hence the MAP estimate can also be seen as the result of a stochastic proximal Bregman point algorithm with step-size $\gamma_n = \inv{n_0 + n}$ at step $n$.

@@

\section{Stochastic Mirror Descent (SMD)}
\subsection{MAP as SMD}
Let $f(\natp) := \expect[x]{-\log p(x|\natp)} = - \langle \mu_*, \natp\rangle + \logpart(\natp)$. In words, $f$ is linear modification of a convex function $\logpart$, which we can access only through noisy estimates of $\mu_*$. It turns out that the MAP estimate can also be seen as the iterates of stochastic mirror descent with mirror map $\logpart$. First let's recall mirror descent iteration
\begin{align}
	\natp_{t+1} 
	&:= \argmin_\natp \lr \linear_f(\natp; \natp_t) + \bregman(\natp ; \natp_t)  \\
	& = \nabla \conj (\nabla \logpart(\natp_t) - \lr \nabla f(\natp_t))
\end{align}
where $\linear_f(\natp; \natp_t) = f(\natp_t) + \langle \nabla f(\natp_t), \natp - \natp_t \rangle$ is the linear approximation of $f$ in $\natp_t$ evaluated at $\natp$. Solving this problem require solving problems of the form $\argmin_\natp - \langle c, \natp \rangle + \logpart(\natp)$, eg computing the convex conjugate of $\logpart$. 
Note that finding $\natp_*$ is done with 1 step of mirror descent. Indeed plugging in definitions of $f$ and $\mu$ yields
\begin{align}
	&\mu_{t+1}  = \mu_t - \lr (\mu_t - \mu_*) \\
	\implies &\mu_t  = \mu_* + (1- \lr)^t (\mu_0 - \mu_*)
\end{align}
which shows exponential convergence and 1-step convergence when $\lr =1$. Back to our sheep, the MAP iteration can be cast as stochastic mirror descent (SMD),  with $g_t = \nabla f(\natp_t, x_{t+1})$ a stochastic estimate  of $\nabla f(\natp_t)$
\begin{align}
    \hat \natp_{n+1} 
    &= \argmin_\natp 
    - \langle T(x_{t+1}), \natp \rangle  + \logpart(\natp) + (n_0 + n) \bregman(\natp ; \natp_n) \\
    %&(n_0+n)\left(\logpart(\natp) - \logpart(\natp_n) - \langle \nabla \logpart(\natp_n)  , \natp - \natp_n \rangle \right)  \\
    &= \argmin_\natp - \langle T(x_{t+1}), \natp \rangle   + \logpart(\natp_n) + \langle \nabla \logpart(\natp_n)  , \natp - \natp_n \rangle + (n_0+n + 1) \bregman(\natp ; \natp_n) \\
    &= \argmin_\natp \linear_f(\natp;\natp_n, x_{t+1}) + (n_0 + n + 1) \bregman(\natp ; \natp_n)
\end{align}
where $\linear_f(\natp;\natp_n, x_{t+1})$ is the stochastic linearization of $f$ at $\natp_n$ evaluated at $\natp$ with randomness coming from $x_{t+1}$. This is the formula for stochastic mirror descent (SMD) applied to $f$ with mirror map $\logpart$ and step-size $\gamma_n = \inv{n_0 + n + 1}$.

\subsection{Relative Smoothness for Mirror Descent}
In the classic setting, SMD is studied under strong-convexity assumption on the mirror map $\logpart$ \citep{bubeck2015convex}.
In our setting this is not always true -- eg gaussians.
However a recent and fast-expanding body of work is concerned with a new assumption: relative smoothness and relative strong-convexity.
\begin{align}
	& \text{$f$ is $L$-smooth relative to $h$} \\
	\iff & Lh - f \ \text{convex} \\
	\iff & f(y) \leq 	f(x) + \langle \nabla f(x) , y-x \rangle + L\cB_h(y;x), \forall x,y \\
	\iff & \cB_f(y;x) \leq L \cB_h(y;x), \forall x,y \\
	\iff & \nabla^2 f(x) \leq L \nabla^2 h(x),  \forall x
\end{align} 
where the last equivalence holds only when $f$ and $h$ are twice differentiable. In words, $f$ is upper bounded by it linear approximation plus the $h$-Bregman divergence, which can also be seen as a bound between divergences, or more locally as a bound between Hessians. Similarly, $f$ is $\mu$ strongly-convex relative to $h$ if
\begin{align}
	& f-\mu h \ \text{ convex} \\
	\iff & f(x) + \langle \nabla f(x) , y-x \rangle + \mu \cB_h(y;x) \leq f(y)\forall x,y \\
	\iff & \mu\cB_h(y;x) \leq \cB_f(y;x) \forall x,y \\
	\iff & \mu \nabla^2 h(x) \leq \nabla^2 f(x)  \forall x \; .
\end{align}

Another way to view this elegant generalization of smoothness and strong-convexity is as a transfer of the Loewner partial order on symmetric matrices to functions, via the Hessian. 
As such it can be applied to many functions that were out of reach for $\ell^2$ norm, by taking the appropriate reference function.  For instance $h(x) = -\log(x)$ or $h(x) = x^4$. 
As early as 2011, \citet{birnbaum2011distributed} showed $O(\inv{t})$ convergence rate for mirror descent under smoothness assumption relative to the mirror map. 
More precisely, he proved that when $f$ is $L$-smooth relative to $h$, then the suboptimality of the sequence 
\begin{align}
	x_{t+1}  = \argmin_x \langle \nabla f(x_t), x \rangle + \cB_h(x ; x_t)
\end{align}
is upper bounded by the simple formula
\begin{align}
	\implies f(x_t) - f(x_*) \leq \frac{L \cB_h(x_* ; x_0) }{t} \; .
\end{align}
These notions were rediscovered and expanded by \citet{bauschke2017descent} and \citet{lu2018relatively}. If you need to read one, pick \citet{lu2018relatively} -- I found it much much easier and more enjoyable to read.  This latter paper also derived a linear convergence rate for mirror descent under relative smoothness and strong-convexity, with the relative condition number $\frac{L}{\mu}$ appearing.

@@

	\expect[X_{1\dots n}]{\bregmanconj ( \mu_* ;\hat \mu) }

@@

Our functions of interest is $f(\natp)= \bregman (\natp ; \natp^*)= \bregmanconj ( \meanp^* ; \meanp) = g(\meanp)$, with minimum $f^*=0$. 
If $\logpart$ is self-concordant, then so is $f$, but not necessarily $g$.
The gradient and Hessian of $f$ are
\begin{align}
    f(\natp) 
    &= \logpart(\natp) - \logpart(\natp^*) - \langle \meanp^*, \natp - \natp^* \rangle \\
    \nabla f(\natp) 
    &= \meanp - \meanp^* = \expect[p(X|\natp)]{T(X)} - \expect[p(X|\natp^*)]{T(X)} \\
    \nabla^2 f(\natp) 
    &= \mSigma(\natp) = \Cov_{p(X|\natp)}[T(X)]
\end{align}
so that we get the bound.
\begin{align}
    \bregmanconj(\mu^* ; \mu) 
    \leq \decrement(\natp)^2 
    = \norm{\mu^* - \mu(\theta)}^2_{\mSigma(\theta)^{-1}} 
    \leq 0.46
\end{align}
% If we were to use $g(\meanp)$ instead, we would get dirty third order derivatives
% \begin{align}
%     g(\meanp) 
%     &= \conj(\meanp^*) - \conj(\meanp) - \langle \natp, \meanp^* - \meanp \rangle \\
%     \nabla g(\meanp) 
%     &= -\natp - \nabla^2 \conj(\meanp) (\meanp^* - \meanp) + \natp \\
%     &= \nabla^2 \conj(\meanp) (\meanp - \meanp^*) \\
%     \nabla^2 g(\meanp) 
%     &= \nabla^2 \conj(\meanp) + \nabla^3 \conj(\meanp ) (\meanp - \meanp^*)
% \end{align}
% where the last expression involves a third order tensor.

Finally, if instead we were looking at a different function switching the role of $\meanp$ and $\meanp^*$,  $h(\meanp) = \bregmanconj ( \meanp ; \meanp^*)$, then we would get
\begin{align}
    \nabla h(\meanp ) 
    &= \natp - \natp^* \\
    \nabla^2 h(\meanp ) 
    &= \nabla^2 \conj(\meanp) 
    = \nabla^2 \logpart(\natp) ^{-1}
    = \Cov_{p(X|\natp)}[T(X)]^{-1}\\
    \implies 
    \decrement(\meanp)
    &= \Var_{p(X|\natp)}[(\natp - \natp^*)^T T(X)] \; .
\end{align}
This is just a remark. I don't think it can help us to get anywhere.

@@

Proofs of convergence for mirror descent with relative smoothness rely on a specific property of Bregman divergences, familiar to information geometry folks. If $x_+$ is solution to 
\begin{align}
	\min	_{x\in C} \overbrace{f(x) + \cB_h(x ; y)}^{\phi(x)}
\end{align}
where $C$ is a closed convex set, $f$ is a convex function, $\cB_h$ is the Bregman divergence induce by some convex function $h$ (we will drop the index, and use a semicolon instead of double bars to lighten the notation), and $y$ is some reference vector. Then 
\begin{align}
	\forall x, f(x) + \cB(x ;y) \geq f(x_+) + \cB(x ; x_+) +\cB(x_+ ; y) \; .
\end{align}
This property is an analog of the Pythagorean theorem for generalized projections (setting $f=0$ and $h=\|.\|^2$).
